\documentclass[11pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{natbib}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{times}
\usepackage{setspace}

% Header and footer setup
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\fancyfoot[C]{\thepage}

% Title page formatting
\title{\vspace{-2cm}\textbf{Laboratory Studies of Cosmology inspired Defect\\Dynamics in Liquid Crystals}}
\author{Anik Mandal$^{1,*}$\\
\\
SLP guide: Prof. Pramoda Kumar$^1$\\
Supervisor: Prof. Somak Raychaudhury$^1$\\
\\
$^1$Ashoka University, Sonipat, India\\
$^*$Email: anik.mandal\_phd24@ashoka.edu.in}
\date{}

\begin{document}

\maketitle
\thispagestyle{empty}

\begin{abstract}
Nematic liquid crystals (NLCs) offer a unique laboratory platform for probing the dynamics of topological defects analogous to cosmic strings, monopoles, and textures formed during symmetry-breaking phase transitions in the early universe. We will induce rapid isotropic–nematic transitions via controlled electric-field and temperature quenches and capture defect dynamics with high-speed polarized-optical videography. String production will be analyzed to validate both the Kibble mechanism of defect creation and its quench-rate prediction (density $\propto \tau_Q^{-1/2}$), while subsequent network coarsening will be quantified against the one-scale annihilation law (density $\propto t^{-1}$). Intercommutation events will be documented to characterize loop formation. In addition, we will explore texture decay—monopole-antimonopole pinch-off—through targeted experiments and one-constant relaxation simulations. This interdisciplinary investigation offers critical laboratory tests of cosmological defect theories and their implications for large-scale structure formation.
\end{abstract}

\chapter{Introduction}

A symmetry-breaking phase transition is a fundamental physical phenomenon where a system undergoes a transition from a more symmetric, higher-energy state to a less symmetric, lower-energy state. During this transition, the system's symmetry present at higher temperatures or energies is spontaneously lost as it reaches a critical point, resulting in an ordered phase characterized by new properties. As the system transitions, different regions may choose different ways to break the symmetry independently, which can cause mismatches where these regions meet. These mismatches create defects, which are places in the material where the order is disrupted or the alignment is not well defined. The dynamical study of defects formed during symmetry-breaking phase transitions is increasingly important in cosmology, particle physics, and condensed matter physics.

The Kibble-Zurek mechanism (KZM) provides a unifying framework to describe defect formation during rapid symmetry-breaking transitions. In cosmology, as the early universe cooled from the Big Bang, the initially unified forces separated through successive symmetry-breaking phase transitions. Kibble proposed \cite{kibble1976topology} that this process occurred unevenly across space, giving rise to mismatches at causal boundaries that manifested as stable topological defects such as domain walls, cosmic strings, and monopoles. These defects are not only relics of the early universe but also form the theoretical basis for several models of cosmic structure formation \cite{turok1989global}.

Zurek extended this cosmological idea to condensed matter systems, suggesting that analogous defects could form in laboratory settings \cite{zurek1985cosmological}. For example, a rapid phase transition in superfluid helium should generate vortex lines as the condensed-matter equivalents of cosmic defects. While practical challenges hindered experimental verification in helium, other systems, such as liquid crystals \cite{gennes1993physics}, provide a more accessible platform. Their phase transitions occur at ordinary temperatures, and the resulting defect dynamics can be directly observed with optical microscopy, making them valuable analogs for probing the Kibble-Zurek mechanism.

\chapter{Project Objectives}

This semester-long project aims to establish nematic liquid crystals (NLCs) as an experimental laboratory system for studying the dynamical evolution of topological defects (strings, monopoles, and textures) that are directly analogous to those predicted in cosmological theories. The primary objective is to experimentally validate key theoretical predictions including the Kibble-Zurek mechanism of defect production during symmetry-breaking phase transitions, the ``one-scale'' scaling model for cosmic string evolution characterized by defect density decay following the annihilation law $\rho \propto t^{-1}$ with time $t$, and string intercommutation processes that occur when cosmic strings cross and reconnect. Additionally, we will test the KZM prediction \cite{fowler2017kibble} $\rho \propto \tau_Q^{-1/2}$, by varying quench rates $\tau_Q$. We also plan to include an investigation of texture decay—generating $\pm 1$ rings in suspended films and using simulations to track monopole–antimonopole formation, subject to project scheduling.

\chapter{Methodology}

The proposed study will build directly upon the experimental framework of Chuang et al. \cite{chuang1991cosmology,chuang1991late} by employing 4-cyano-4'-n-pentylbiphenyl (5CB) nematic liquid crystal as an accessible analog for cosmological defects and precisely replicating their temperature quench apparatus to induce rapid isotropic-to-nematic transitions. Samples will be confined within a custom cell maintained at 33$^\circ$C, will generate dense networks of strings and monopoles via the Kibble mechanism. A high-speed optical microscope will record the resulting defect tangle over intervals from 1 to 32 s, allowing automated image analysis to extract string density $\rho(t)$.

To test the Kibble–Zurek prediction of $\rho \propto \tau_Q^{-1/2}$ scaling law, we will perform controlled electric-field ``quenches'' in NLC cells by varying the field-ramp time $\tau_Q$. To examine texture decay, $\pm 1$ defect loops will be hand-drawn in freely suspended films using a micrometer-driven glass probe and their monopole-antimonopole pinch-off tracked under optical microscopy; simple relaxation simulations on a $40^3$ lattice in the one-constant approximation will replicate this process.

\chapter{Expected Outcomes}

\begin{itemize}
\item We are expecting to find direct visualization of monopoles and strings in nematic liquid crystals, confirming their predicted structures and behaviors under unequal elastic constants.

\item We are expecting to validate the one-scale scaling model for string network evolution, with string density following a $\rho \propto t^{-1}$ power law.

\item We are expecting to document string intercommutation events, supporting the reconnection processes important for loop formation and network evolution.

\item We are expecting to validate the Kibble-Zurek scaling law $\rho \propto \tau_Q^{-1/2}$, demonstrating how defect density depends on the quench rate during phase transitions.

\item We are expecting to observe texture decay through the formation and annihilation of monopole–antimonopole pairs, providing the laboratory evidence of texture unwinding consistent with simulations.

\item We are expecting to measure texture decay rates and lifetimes that explain the rarity of stable textures.
\end{itemize}

These outcomes will demonstrate that nematic liquid crystals serve as effective experimental analogs for cosmological defect phenomena, offering critical tests of the Kibble mechanism, scaling laws, and topological defect decay relevant to large-scale structure formation.

\bibliographystyle{unsrt}
\begin{thebibliography}{9}

\bibitem{chuang1991cosmology}
Chuang, I., Durrer, R., Turok, N. \& Yurke, B. 
Cosmology in the Laboratory: Defect Dynamics in Liquid Crystals. 
\textit{Science} \textbf{251}, 1336–1342 (1991). 
URL \url{https://doi.org/10.1126/science.251.4999.1336}.

\bibitem{chuang1991late}
Chuang, I., Turok, N. \& Yurke, B. 
Late-time coarsening dynamics in a nematic liquid crystal. 
\textit{Physical Review Letters} \textbf{66}, 2472 (1991). 
URL \url{https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.66.2472}.

\bibitem{kibble2007phase}
Kibble, T. 
Phase-transition dynamics in the lab and the universe. 
\textit{Physics Today} \textbf{60}, 47–52 (2007). 
URL \url{https://doi.org/10.1063/1.2784684}.

\bibitem{kibble1976topology}
Kibble, T. W. 
Topology of cosmic domains and strings. 
\textit{Journal of Physics A: Mathematical and General} \textbf{9}, 1387 (1976). 
URL \url{https://iopscience.iop.org/article/10.1088/0305-4470/9/8/029}.

\bibitem{turok1989global}
Turok, N. 
Global texture as the origin of cosmic structure. 
\textit{Physical Review Letters} \textbf{63}, 2625 (1989). 
URL \url{https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.63.2625}.

\bibitem{zurek1985cosmological}
Zurek, W. H. 
Cosmological experiments in superfluid helium? 
\textit{Nature} \textbf{317}, 505–508 (1985). 
URL \url{https://www.nature.com/articles/317505a0}.

\bibitem{gennes1993physics}
Gennes, P. G. D. \& Prost, J. 
\textit{The Physics of Liquid Crystals}. 
The Physics of Liquid Crystals (1993). 
URL \url{https://academic.oup.com/book/54859}.

\bibitem{fowler2017kibble}
Fowler, N. \& Dierking, D. I. 
Kibble–Zurek Scaling during Defect Formation in a Nematic Liquid Crystal. 
\textit{ChemPhysChem} \textbf{18}, 812–816 (2017). 
URL \url{https://doi.org/10.1002/cphc.201700023}.

\end{thebibliography}

\end{document}